Supplementary material: The increasing cost of happiness

### 2021-10-27 src/supplementary-SSM2.Rmd
Final response to Reviewer 1. Closes issue #27

##### Figure S2. Conditional model parameters  
```{r figure_s2, fig.width = 6, fig.height = 8, dpi=300}
nl_fits = read_rds("../results/gh9_hhoecd_0cov_tight.RDS")

nl_fits_5cov = read_rds("../results/gh9_hhoecd_5cov_tight.RDS")

income_params <- read_rds("../results/gh9_hhoecd_summary.RDS") %>%
  bind_rows(.id = "fit")

posteriors_5c <- map_dfr(nl_fits_5cov, as.data.frame, .id = "fit") %>%
  left_join(income_params) %>%
  mutate(omega_dollars = (b_omega_Intercept * sd + mean)/1000) %>%
  as_tibble()

param_key = c(
  omega_dollars = "Change point ($000s)",
  b_b0_Intercept = paste0("Intercept (\u03b2", "0)"),
  b_b1_Intercept = paste0("Pre-slope (\u03b2", "1)"),
  b_b2_Intercept = paste0("Post-slope (\u03b2", "2)")
)

posteriors_5c %>%
  select(fit, contains("b0"), contains("b1"), contains("b2"), omega_dollars) %>%
  gather(key = "param", val = "sample", -fit) %>%
  mutate(
    year = extract_numeric(fit),
    param = recode_factor(param, !!!param_key)
  ) %>% 
  group_by(param, year) %>%
  summarise(
    low = quantile(sample, probs = 0.05, na.rm=T),
    expected = quantile(sample, probs = 0.5, na.rm=T),
    upp = quantile(sample, probs = 0.95, na.rm=T)
  ) %>%
  mutate(model = "conditional") -> parameter_summary_5c

posteriors_0c <- map_dfr(nl_fits, as.data.frame, .id = "fit") %>%
  left_join(income_params) %>%
  mutate(omega_dollars = (b_omega_Intercept * sd + mean)/1000) %>%
  as_tibble()

posteriors_0c %>%
  select(fit, contains("b0"), contains("b1"), contains("b2"), omega_dollars) %>%
  gather(key = "param", val = "sample", -fit) %>%
  mutate(
    year = extract_numeric(fit),
    param = recode_factor(param, !!!param_key)
  ) %>% 
  group_by(param, year) %>%
  summarise(
    low = quantile(sample, probs = 0.05, na.rm=T),
    expected = quantile(sample, probs = 0.5, na.rm=T),
    upp = quantile(sample, probs = 0.95, na.rm=T)
  ) %>%
  mutate(model = "unconditional") -> parameter_summary_0c

bind_rows(parameter_summary_0c, parameter_summary_5c) %>%
  ggplot(aes(y = year, x = expected, color = model)) +
  geom_pointrange(aes(xmin = low, xmax = upp), position = position_dodge(width = 0.75)) +
  scale_y_continuous(breaks = c(2001, 2007, 2013, 2019)) +
  facet_wrap(~param, scales = "free_x") +
  ft_theme(base_size = 12) +
  scale_color_manual(values = c("#C79999", "grey65")) +
  theme(legend.position = "none",
        axis.title = element_blank())
```
```
Figure legend: Comparison of the conditional parameter estimates (red) and unconditional parameter estimates (grey) in the piecewise-linear model of income on affective wellbeing. The conditional models included covariates for age, age squared, sex, education, and illness. Horizontal bar represents the 95% credible region and the solid point indicates the expected value (median) of each distribution.
```

@@

Household income values for high income individuals are substituted with the average income value of all those above a certain threshold (topcoded). This is performed by the University of Melbourne to preserve anonymity, and we excluded people marked as topcoded in the main analysis (highest _n_ in any year was _n_ = 103 in 2019). Here we present the parameter estimates of the affective wellbeing models with topcodes included, and compare them with parameters from models which exclude them.

##### Figure S3. Topcoded model parameters  
```{r figure_s3, fig.width = 6, fig.height = 8, dpi=300}
nl_fits_topcodes = read_rds("../results/fits/inc_topcodes/gh9_hhoecd_0cov_tight.RDS")

topcode_income_params <- read_rds("../results/fits/inc_topcodes/gh9_hhoecd_data.RDS") %>%
  bind_rows(.id = "fit")

posteriors_topcodes <- map_dfr(nl_fits_topcodes, as.data.frame, .id = "fit") %>%
  left_join(topcode_income_params) %>%
  mutate(omega_dollars = (b_omega_Intercept * sd + mean)/1000) %>%
  as_tibble()

posteriors_topcodes %>%
  select(fit, contains("b0"), contains("b1"), contains("b2"), omega_dollars) %>%
  gather(key = "param", val = "sample", -fit) %>%
  mutate(
    year = extract_numeric(fit),
    param = recode_factor(param, !!!param_key)
  ) %>% 
  group_by(param, year) %>%
  summarise(
    low = quantile(sample, probs = 0.05, na.rm=T),
    expected = quantile(sample, probs = 0.5, na.rm=T),
    upp = quantile(sample, probs = 0.95, na.rm=T)
  ) %>%
  mutate(model = "topcodes") -> parameter_summary_topcodes

parameter_summary_0c %>%
  mutate(model = "no topcodes") -> parameter_summary_excl

bind_rows(parameter_summary_excl, parameter_summary_topcodes) %>%
  ggplot(aes(y = year, x = expected, color = model)) +
  geom_pointrange(aes(xmin = low, xmax = upp), position = position_dodge(width = 0.75)) +
  scale_y_continuous(breaks = c(2001, 2007, 2013, 2019)) +
  facet_wrap(~param, scales = "free_x") +
  ft_theme(base_size = 12) +
  scale_color_manual(values = c("grey65", "#C79999")) +
  theme(legend.position = "none",
        axis.title = element_blank())
```
```
Figure legend: Comparison of the topcoded parameter estimates (red) and parameter estimates excluding topcodes (grey) in the piecewise-linear model of income on affective wellbeing. The topcoded models included n = 26 to 130 extra (high income) individuals in each year. Horizontal bar represents the 95% credible region and the solid point indicates the expected value (median) of each distribution.
```

@@

Household income values were adjusted for household size in the main analysis. The _maximum_ household sizes varied from 10 (2017, 2019) to as high as 17 (2016), and so adjustment accounts for larger households needing more resources to achieve the same standard of living as a smaller household, and for some ‘economies of scale’ when sharing living costs. For example, two people living alone in two separate households typically incur higher living costs in total than two people living in one household. Here we present the parameter estimates of the affective wellbeing models without adjustment for household size and with adjustment.

@@

posteriors_unadjsize <- map_dfr(nl_fits_unadjsize, as.data.frame, .id = "fit") %>%
  left_join(income_params) %>%
  mutate(omega_dollars = (b_omega_Intercept * sd + mean)/1000) %>%
  as_tibble()

posteriors_unadjsize %>%
  select(fit, contains("b0"), contains("b1"), contains("b2"), omega_dollars) %>%
  gather(key = "param", val = "sample", -fit) %>%
  mutate(
    year = extract_numeric(fit),
    param = recode_factor(param, !!!param_key)
  ) %>% 
  group_by(param, year) %>%
  summarise(
    low = quantile(sample, probs = 0.05, na.rm=T),
    expected = quantile(sample, probs = 0.5, na.rm=T),
    upp = quantile(sample, probs = 0.95, na.rm=T)
  ) %>%
  mutate(model = "unadjusted") -> parameter_summary_unadjsize

@@

bind_rows(parameter_summary_unadjsize, parameter_summary_adj) %>%
  ggplot(aes(y = year, x = expected, color = model)) +
  geom_pointrange(aes(xmin = low, xmax = upp), position = position_dodge(width = 0.75)) +
  scale_y_continuous(breaks = c(2001, 2007, 2013, 2019)) +
  facet_wrap(~param, scales = "free_x") +
  ft_theme(base_size = 12) +
  scale_color_manual(values = c("grey65", "#C79999")) +
  theme(legend.position = "none",
        axis.title = element_blank())
```
```
Figure legend: Comparison of the parameter estimates unadjusted for household size (red) and parameter estimates adjusted for size (grey) in the piecewise-linear model of income on affective wellbeing. Horizontal bar represents the 95% credible region and the solid point indicates the expected value (median) of each distribution.
```

@@

income_params <- read_rds("../results/gh9_hhoecd_summary.RDS") %>%
  bind_rows(.id = "fit")

posteriors_pos <- map_dfr(gh9pos.fits, as.data.frame, .id = "fit") %>%
  left_join(income_params) %>%
  mutate(omega_dollars = (b_omega_Intercept * sd + mean)/1000) %>%
  as_tibble()

param_key = c(
  omega_dollars = "Change point ($000s)",
  b_b0_Intercept = paste0("Intercept (\u03b2", "0)"),
  b_b1_Intercept = paste0("Pre-slope (\u03b2", "1)"),
  b_b2_Intercept = paste0("Post-slope (\u03b2", "2)")
)

posteriors_pos %>%
  select(fit, contains("b0"), contains("b1"), contains("b2"), omega_dollars) %>%
  gather(key = "param", val = "sample", -fit) %>%
  mutate(
    year = extract_numeric(fit),
    param = recode_factor(param, !!!param_key)
  ) %>% 
  group_by(param, year) %>%
  summarise(
    low = quantile(sample, probs = 0.05, na.rm=T),
    expected = quantile(sample, probs = 0.5, na.rm=T),
    upp = quantile(sample, probs = 0.95, na.rm=T)
  ) %>%
  mutate(model = "positive") -> parameter_summary_pos

posteriors_neg <- map_dfr(gh9neg.fits, as.data.frame, .id = "fit") %>%
  left_join(income_params) %>%
  mutate(omega_dollars = (b_omega_Intercept * sd + mean)/1000) %>%
  as_tibble()

posteriors_neg %>%
  select(fit, contains("b0"), contains("b1"), contains("b2"), omega_dollars) %>%
  gather(key = "param", val = "sample", -fit) %>%
  mutate(
    year = extract_numeric(fit),
    param = recode_factor(param, !!!param_key)
  ) %>% 
  group_by(param, year) %>%
  summarise(
    low = quantile(sample, probs = 0.05, na.rm=T),
    expected = quantile(sample, probs = 0.5, na.rm=T),
    upp = quantile(sample, probs = 0.95, na.rm=T)
  ) %>%
  mutate(model = "negative") -> parameter_summary_neg

@@

#### Parameter and fit summary statistics

@@

For all models estimated with RStan presented in this report we confirmed the absence of divergent transitions; the $\hat{R}$ approached 1 indicating convergance; and the effective sample size (ESS) for each parameter was _n_ > 300, which is a conservative threshold for estimation.

Table S2 below shows the fit and summary statistics for each piecewise-linear model fit of affective wellbeing and income in each year. Estimates for the intercept, pre-slope and post-slope are in year-wise SD units while the change-point is presented in real 2019 dollars.

##### Table S2. Parameter estimates and fit statistics for piecewise-linear fit of affective wellbeing and income
```{r table_s2, ft.align="left"}
nl_fits = read_rds("../results/gh9_hhoecd_0cov_tight.RDS")

income_params %>%
  mutate(fit = str_remove(fit, "fit")) %>%
  column_to_rownames("fit") -> income

param_key = c(
  omega_Intercept = "Change point ($000s)",
  b0_Intercept = paste0("Intercept (\u03b2", "0)"),
  b1_Intercept = paste0("Pre-slope (\u03b2", "1)"),
  b2_Intercept = paste0("Post-slope (\u03b2", "2)")
)

# param_key = c(
#   b0_Intercept = "Intercept",
#   b1_Intercept = "Pre-slope",
#   b2_Intercept = "Post-slope",
#   omega_Intercept = "Change-point ($000s)"
# )

map_dfr(nl_fits, 
        .f = ~{
          fit_summary <- summary(.x, prob = 0.95)

@@

          fit_summary$fixed %>%
            as_tibble(rownames = "Parameter") %>%
            bind_rows(sigma_summary) %>%
            select(Parameter, everything())

          },
        .id = "Year") %>%
  mutate(
    Year = str_remove(Year, "fit"),
    Parameter = recode_factor(Parameter, !!!param_key),
    Estimate = if_else(str_detect(Parameter, "Change"),
                       (Estimate * income[Year, "sd"] + income[Year, "mean"])/1000,
                       Estimate),
    Est.Error = if_else(str_detect(Parameter, "Change"),
                        (Est.Error * income[Year, "sd"])/1000,
                        Est.Error),
    `l-95% CI` = if_else(str_detect(Parameter, "Change"),
                        (`l-95% CI` * income[Year, "sd"] + income[Year, "mean"])/1000,
                        `l-95% CI`),
    `u-95% CI` = if_else(str_detect(Parameter, "Change"),
                         (`u-95% CI` * income[Year, "sd"] + income[Year, "mean"])/1000,
                         `u-95% CI`)
    ) %>%
  flextable() %>%
  merge_v(j = 1) %>%
  valign(j = 1, valign = "top") %>%
  bg(i = ~Year %in% seq(2001, 2019, 2), 
     bg = "grey90", 
     part = "body") %>%
  colformat_double(j = 3:6, digits= 3) %>%
  colformat_double(j = 7, digits= 1) %>%
  footnote(i = 1, j = 7,
           value = as_paragraph(
             c("Rhat is the potential scale reduction factor on split chains")),
           ref_symbols = c("a"),
           part = "header") %>%
  footnote(i = 1, j = c(8:9),
           value = as_paragraph(
             c("Bulk_ESS and Tail_ESS are effective sample size (ESS) measures")),
           ref_symbols = c("b"),
           part = "header") %>%
  fontsize(size = 9, part = "all") %>%
  width(j = 1, width = .5) %>%
  width(j = 2, width = 1) %>%
  width(j = c(3:7), width = .7) %>%
  width(j = c(8, 9), width = .8)
```

@@

The coeficients (in SD units) for the piecewise-linear, log-linear and linear Bayesian models are shown in the table below.

##### Table S3. Bayesian coefficients from linear, log-linear and piecewise-linear models  
```{r table_s3}
map_df(log_fits, .f = ~{
  fixef(.x) %>%
    as_tibble(rownames = "terms") %>%
    transmute(terms, coef = paste0(round(Estimate, 2), " (", round(Est.Error, 3), ")")) %>%
    spread(terms, coef)},
  .id = "Year"
  ) %>%
  rename("b0_log" = 2, "b1_log" = 3) -> coefs_gh9_log

map_df(lm_fits, .f = ~{
  fixef(.x) %>%
    as_tibble(rownames = "terms") %>%
    transmute(terms, coef = paste0(round(Estimate, 2), " (", round(Est.Error, 3), ")")) %>%
    spread(terms, coef)},
  .id = "Year"
  ) %>%
  select(Year, "b0_lin" = 3, "b1_lin" = 2) -> coefs_gh9_lm

map_df(nl_fits, .f = ~{
  fixef(.x) %>%
    as_tibble(rownames = "terms") %>%
    transmute(terms, coef = paste0(round(Estimate, 2), " (", round(Est.Error, 3), ")")) %>%
    spread(terms, coef)},
  .id = "Year"
  ) %>%
  rename("b0_pw" = 2, "b1_pw" = 3, 
         "b2_pw" = 4, "Omega" = 5) -> coefs_gh9_pw

coefs_affective <- left_join(coefs_gh9_pw, coefs_gh9_log) %>%
  left_join(coefs_gh9_lm) %>%
  mutate(Year = str_remove(Year, "fit"),
         Wellbeing = "Affective")

map_df(log_fits_losat, .f = ~{
  fixef(.x) %>%
    as_tibble(rownames = "terms") %>%
    transmute(terms, coef = paste0(round(Estimate, 2), " (", round(Est.Error, 3), ")")) %>%
    spread(terms, coef)},
  .id = "Year"
) %>%
  rename("b0_log" = 2, "b1_log" = 3) -> coefs_losat_log

map_df(lm_fits_losat, .f = ~{
  fixef(.x) %>%
    as_tibble(rownames = "terms") %>%
    transmute(terms, coef = paste0(round(Estimate, 2), " (", round(Est.Error, 3), ")")) %>%
    spread(terms, coef)},
  .id = "Year"
) %>%
  select(Year, "b0_lin" = 3, "b1_lin" = 2) -> coefs_losat_lm

map_df(nl_fits_losat, .f = ~{
  fixef(.x) %>%
    as_tibble(rownames = "terms") %>%
    transmute(terms, coef = paste0(round(Estimate, 2), " (", round(Est.Error, 3), ")")) %>%
    spread(terms, coef)},
  .id = "Year"
) %>%
  rename("b0_pw" = 2, "b1_pw" = 3, 
         "b2_pw" = 4, "Omega" = 5) -> coefs_losat_pw

coefs_cognitive <- left_join(coefs_losat_pw, coefs_losat_log) %>%
  left_join(coefs_losat_lm) %>%
  mutate(Year = str_remove(Year, "fit"),
         Wellbeing = "Cognitive")

bind_rows(coefs_affective, coefs_cognitive) %>%
  select(Wellbeing, everything()) %>%
  flextable() %>%
  merge_v(j = 1) %>%
  set_header_labels(values = c(b0_pw = paste0("\u03b2", "0"),
                               b1_pw = paste0("\u03b2", "1"),
                               b2_pw = paste0("\u03b2", "2"),
                               Omega = "\u03C9",
                               b0_log = paste0("\u03b2", "0"),
                               b1_log = paste0("\u03b2", "1"),
                               b0_lin = paste0("\u03b2", "0"),
                               b1_lin = paste0("\u03b2", "1"))) %>%
  add_header(top = TRUE,
             values = c(b0_pw = "Piecewise-linear (se)",
                        b1_pw = "Piecewise-linear (se)",
                        b2_pw = "Piecewise-linear (se)",
                        Omega = "Piecewise-linear (se)",
                        b0_log = "Log-linear (se)",
                        b1_log = "Log-linear (se)",
                        b0_lin = "Linear (se)",
                        b1_lin = "Linear (se)")) %>%
  merge_h(part = "header") %>%
  width(j = 1:2, width = c(0.75, 0.5)) %>%
  width(j = 3:10, width = 0.65) %>%
  fontsize(size = 9, part = "all") %>%
  # flextable::theme_vanilla() %>% # theme_vanilla() 
  valign(valign = "top") %>%
  valign(valign = "bottom", part = "header") %>%
  border_remove() %>%
  hline_top(part="header", border = officer::fp_border()) %>%
  hline_bottom(part="header", border = officer::fp_border()) %>%
  hline(i = 19, border = officer::fp_border()) %>%
  hline_bottom(part="body", border = officer::fp_border())
```

The table of model parameters (untransformed) for each year show that the $\beta_1$ slopes for piecewise-linear, log-linear and linear model for affective wellbeing and income do not show a sustained trend ove the 19 year period. The $\beta_1$ coefficients increase in the initial 5-10 years and then decline. Despite this common pattern, the linear slope appears to show the most sustained increase while the log-linear slope shows a relatively less sustained increase. Overall, the pattern of changes among the models indicates the function between income and affective wellbeing has in fact become slightly steeper over the period.

#### Prior and residual checks

@@

The posterior estimates of the piecewise model indicated that the relationship between affective wellbeing and income changed over time between 2001 and 2019. One implication of such a change is the disparity in affective wellbeing between income groups has increased. This will occur as more people fall below the change point over time and so are subject to the steep region of the function where income and affective wellbeing are strongly related. To directly examine the implications of our model for such inequities in affective wellbeing, we used the model to estimate or _predict_ the affective wellbeing of two different income levels: one income level which was above the change point in 2001 but fell below it by 2019 (\$50K/yr); and another level which always remained above the change points over the same period (\$75K/yr). The figure below presents the affective wellbeing levels (in SD units/year) for each income level as well as the difference in affective wellbeing between them (∆).

@@

##### Figure S8. affective wellbeing (SD units) at $50K/yr and $75K/yr from 2001-2019  
```{r figure_s8, fig.width=7, fig.height=5, dpi=300}
nl_fits = read_rds("../results/gh9_hhoecd_0cov_tight.RDS")

# Optional
# nl_fits_5cov = read_rds(
#   paste0(here, "results/fits/excl_topcodes/sd/gh9_hhoecd_5cov_tight.RDS")
# )

fixed_income_hash <- read_rds("../results/gh9_hhoecd_summary.RDS") %>%
  bind_rows(.id = "fit") %>%
  mutate(low = (40000 - mean)/sd,
         high = (75000 - mean)/sd) %>%
  column_to_rownames(var = "fit") %>%
  as.matrix()

@@

fixed_income_fit %>%
  mutate(year = extract_numeric(fit)) %>%
  select(Estimate, Dollars, year) %>%
  filter(Dollars %in% c("$75K", "$50K")) %>%
  spread(Dollars, Estimate) %>%
  mutate(`$75K - $50K (∆)` = `$75K` - `$50K`) %>%
  gather(income, happiness, -year) %>%
  mutate(income = fct_relevel(income, "$50K", "$75K")) %>%
  ggplot(aes(x = year, y = happiness, color = income)) +
  geom_point(size = 4, alpha = 0.25) +
  geom_smooth(method = "loess", se = F, span = 3) +
  facet_wrap(~income) +
  labs(caption = "Source:     HILDA, University of Melbourne
                 Predicted affective wellbeing scores (SD unit) from piecewise regression on real household income", 
       x = "", y = "") -> p

p +
  ft_theme() +
  ggthemes::scale_colour_wsj() +
  theme(legend.position = "none",
        plot.title = element_text(size = 19),
        strip.text = element_text(size = 12))
```

@@

The figure above shows the disparity in affective wellbeing between two fixed income levels, \$50K/year (traversing the change points) and \$75K/year (above the change points). In 2001, a household income of \$50K/year achieved an above average level of affective wellbeing relative to everyone else that year, however by 2019 affective wellbeing had declined to lower than average levels relative to the population. By contrast, a household income of \$75K/year enjoyed a higher than average level of affective wellbeing for the entire period. The difference (∆) in affective wellbeing between these two income levels doubled over the period. The increasing disparity in affective wellbeing between the rich and the poor implies that affective wellbeing has became more inequitable over time.

@@

#### OLS Insample fit statistics

>>>>

#### Sensitivity tests of model parameters

```{r std_params}
std_params <- list()

for (i in c(2001:2019)) {

  hilda_data %>%
    filter(
      year == i,
      !student,
      hh_disp_inc_eq >= 0,
      !top_hifdip
    ) %>%
    transmute(
      sgh9 = scale(gh9),
      shh_disp_inc_eq = scale(hh_disp_inc_eq),
    ) %>% 
    na.omit() -> df

  income_sd_year <- attr(df$shh_disp_inc_eq, "scaled:scale")
  income_m_year <- attr(df$shh_disp_inc_eq, "scaled:center")
  gh9_sd_year <- attr(df$sgh9, "scaled:scale")
  gh9_m_year <- attr(df$sgh9, "scaled:center")

  std_params[[paste0("fit", i)]] <- c(`mean_income` = income_m_year, 
                       `sd_income` = income_sd_year,
                       `mean_gh9` = gh9_m_year,
                       `sd_gh9` = gh9_sd_year)

std_params <- bind_rows(std_params, .id = "fit")
```

##### Figure S2. Conditional model parameters  
```{r figure_s2, fig.width = 6, fig.height = 8, dpi=300}
nl_fits = read_rds("../results/gh9_hhoecd_0cov_tight.RDS")
nl_fits_5cov = read_rds("../results/gh9_hhoecd_5cov_tight.RDS")

summarise_posteriors <- function(.fits, .params = std_params, .model = NULL) {

  posteriors <- map_dfr(.fits, as.data.frame, .id = "fit") %>%
    left_join(.params, by = "fit") %>%
    transmute(
      fit,
      omega_dollars = (b_omega_Intercept * sd_income + mean_income)/1000,
      b0 = b_b0_Intercept * sd_gh9 + mean_gh9,
      b1 = (b_b1_Intercept * sd_gh9)/10, # change per $1000
      b2 = (b_b2_Intercept * sd_gh9)/10  # change per $1000
    ) %>%
    as_tibble()

  param_key = c(
    omega_dollars = "Change point ($000s)",
    b0 = paste0("Intercept (\u03b2", "0)"),
    b1 = paste0("Pre-slope (\u03b2", "1)"),
    b2 = paste0("Post-slope (\u03b2", "2)")
  )

  posteriors %>%
    gather(key = "param", val = "sample", -fit) %>%
    mutate(
      year = extract_numeric(fit),
      param = recode_factor(param, !!!param_key)
    ) %>% 
    group_by(param, year) %>%
    summarise(
      low = quantile(sample, probs = 0.05, na.rm=T),
      expected = quantile(sample, probs = 0.5, na.rm=T),
      upp = quantile(sample, probs = 0.95, na.rm=T)
    ) %>%
    mutate(model = .model) -> parameter_summary

  return(parameter_summary)

parameter_summary_5c <- summarise_posteriors(.fits = nl_fits_5cov, 
                                             .model = "conditional")

parameter_summary_0c <- summarise_posteriors(.fits = nl_fits,
                                             .model = "unconditional")

bind_rows(parameter_summary_0c, parameter_summary_5c) %>%
  ggplot(aes(y = year, x = expected, color = model)) +
  geom_pointrange(aes(xmin = low, xmax = upp), position = position_dodge(width = 0.75)) +
  scale_y_continuous(breaks = c(2001, 2007, 2013, 2019)) +
  facet_wrap(~param, scales = "free_x") +
  ft_theme(base_size = 12) +
  scale_color_manual(values = c("#A25050", "grey65")) +
  theme(legend.position = "none",
        axis.title = element_blank())
```
```
Figure legend: Comparison of the conditional parameter estimates (red) and unconditional parameter estimates (grey) in the piecewise-linear model of income on affective wellbeing. The conditional models included covariates for age, age squared, sex, education, and illness. Horizontal bar represents the 95% credible region and the solid point indicates the expected value (median) of each distribution.
```

@@

Household income values for high income individuals are substituted with the average income value of all those above a certain threshold (topcoded). This is performed by the University of Melbourne to preserve anonymity, and we excluded people marked as topcoded in the main analysis (highest _n_ in any year was _n_ = 103 in 2019). Here we present the parameter estimates of the affective wellbeing models with topcodes included, and compare them with parameters from models which exclude them (as in the main text).

##### Figure S3. Model parameters including topcoded individuals 
```{r figure_s3, fig.width = 6, fig.height = 8, dpi=300}
nl_fits_topcodes = read_rds("../results/fits/inc_topcodes/gh9_hhoecd_0cov_tight.RDS")

# topcode_income_params <- read_rds("../results/fits/inc_topcodes/gh9_hhoecd_data.RDS") %>%
#   bind_rows(.id = "fit")

topcode_params <- list()

for (i in c(2001:2019)) {

  hilda_data %>%
    filter(
      year == i,
      !student,
      hh_disp_inc_eq >= 0
    ) %>%
    transmute(
      sgh9 = scale(gh9),
      shh_disp_inc_eq = scale(hh_disp_inc_eq),
    ) %>% 
    na.omit() -> df

  income_sd_year <- attr(df$shh_disp_inc_eq, "scaled:scale")
  income_m_year <- attr(df$shh_disp_inc_eq, "scaled:center")
  gh9_sd_year <- attr(df$sgh9, "scaled:scale")
  gh9_m_year <- attr(df$sgh9, "scaled:center")

  topcode_params[[paste0("fit", i)]] <- c(`mean_income` = income_m_year, 
                       `sd_income` = income_sd_year,
                       `mean_gh9` = gh9_m_year,
                       `sd_gh9` = gh9_sd_year)

topcode_params <- bind_rows(topcode_params, .id = "fit")

parameter_summary_topcodes <- summarise_posteriors(nl_fits_topcodes,
                                                   .params = topcode_params,
                                                   .model = "topcodes")

parameter_summary_excl <- parameter_summary_0c %>% mutate(model = "no topcodes")

bind_rows(parameter_summary_excl, parameter_summary_topcodes) %>%
  ggplot(aes(y = year, x = expected, color = model)) +
  geom_pointrange(aes(xmin = low, xmax = upp), position = position_dodge(width = 0.75)) +
  scale_y_continuous(breaks = c(2001, 2007, 2013, 2019)) +
  facet_wrap(~param, scales = "free_x") +
  ft_theme(base_size = 12) +
  scale_color_manual(values = c("grey65", "#A25050")) +
  theme(legend.position = "none",
        axis.title = element_blank())
```
```
Figure legend: Comparison of the topcoded parameter estimates (red) and parameter estimates excluding topcodes (grey) in the piecewise-linear model of income on affective wellbeing. The topcoded models included n = 26 to 130 extra (high income) individuals in each year. Horizontal bar represents the 95% credible region and the solid point indicates the expected value (median) of each distribution.
```

@@

Household income values were adjusted for household size in the main analysis. The _maximum_ household sizes varied from 10 (2017, 2019) to as high as 17 (2016), and so adjustment accounts for larger households needing more resources to achieve the same standard of living as a smaller household, and for some ‘economies of scale’ when sharing living costs. For example, two people living alone in two separate households typically incur higher living costs in total than two people living in one household. Here we present the parameter estimates of the affective wellbeing models without adjustment for household size, and with adjustment (as in the main text).

@@

parameter_summary_unadjsize <- summarise_posteriors(nl_fits_unadjsize, 
                                                    .model = "unadjusted")

@@

bind_rows(parameter_summary_unadjsize, parameter_summary_adj) %>%
  ggplot(aes(y = year, x = expected, color = model)) +
  geom_pointrange(aes(xmin = low, xmax = upp), position = position_dodge(width = 0.75)) +
  scale_y_continuous(breaks = c(2001, 2007, 2013, 2019)) +
  facet_wrap(~param, scales = "free_x") +
  ft_theme(base_size = 12) +
  scale_color_manual(values = c("grey65", "#A25050")) +
  theme(legend.position = "none",
        axis.title = element_blank())
```
```
Figure legend: Comparison of the parameter estimates unadjusted for household size (red) and parameter estimates adjusted for size (grey) in the piecewise-linear model of income on affective wellbeing. Horizontal bar represents the 95% credible region and the solid point indicates the expected value (median) of each distribution.
```

@@

parameter_summary_pos <- summarise_posteriors(gh9pos.fits, .model = "positive")
parameter_summary_neg <- summarise_posteriors(gh9neg.fits, .model = "negative")

@@

#### Sampling and fit statistics

Below we provide sampling statistics (e.g., $\hat{R}$), as well as prior predictions and residual plots.

@@

For all models estimated with RStan presented in this report we confirmed the absence of divergent transitions; the $\hat{R}$ for each parameter approached 1 indicating convergance; and the effective sample size (ESS) for each parameter was _n_ > 300, which is a conservative minimum threshold for estimation.

Table S2 below shows the sampling statistics for each piecewise-linear model fit of affective wellbeing and income in each year.

##### Table S2. Sampling statistics for piecewise-linear fit of affective wellbeing and income
```{r table_s2, ft.align="left"}
nl_fits = read_rds("../results/gh9_hhoecd_0cov_tight.RDS")

params_df <- std_params %>% 
  mutate(fit = str_remove(fit, "fit")) %>%
  column_to_rownames("fit")

map_dfr(nl_fits, 
        .f = ~{
          fit_summary <- summary(.x)

@@

          fit_summary$fixed %>%
            as_tibble(rownames = "Parameter") %>%
            bind_rows(sigma_summary) %>%
            select(Parameter, Bulk_ESS, Tail_ESS)

          },
        .id = "Year") %>%
  mutate(
    Year = str_remove(Year, "fit"),
  ) %>%
  pivot_wider(names_from = Parameter, 
              values_from = c(Bulk_ESS, Tail_ESS)
  ) %>% 
  select(-one_of(c("Bulk_ESS_sigma", "Tail_ESS_sigma"))) %>%
  flextable() %>%
  set_header_labels(
    values = c(Bulk_ESS_b0_Intercept = paste0("\u03b2", "0"),
               Bulk_ESS_b1_Intercept = paste0("\u03b2", "1"),
               Bulk_ESS_b2_Intercept = paste0("\u03b2", "2"),
               Bulk_ESS_omega_Intercept = "\u03C9",
               Tail_ESS_b0_Intercept = paste0("\u03b2", "0"),
               Tail_ESS_b1_Intercept = paste0("\u03b2", "1"),
               Tail_ESS_b2_Intercept = paste0("\u03b2", "2"),
               Tail_ESS_omega_Intercept = "\u03C9")
    ) %>%
  add_header(
    top = TRUE,
    values = c(Bulk_ESS_b2_Intercept = "Bulk ESS",
               Tail_ESS_b2_Intercept = "Tail ESS")) %>%
  merge_h(part = "header") %>%
  width(j = 1, width = 0.5) %>%
  width(j = 2:9, width = 0.625) %>%
  fontsize(size = 9, part = "all") %>%
  valign(valign = "top") %>%
  valign(valign = "bottom", part = "header") %>%
  border_remove() %>%
  hline_top(part="header", border = officer::fp_border()) %>%
  hline_bottom(part="header", border = officer::fp_border()) %>%
  hline_bottom(part="body", border = officer::fp_border())
```

@@

@@

The posterior estimates of the piecewise model indicated that the relationship between affective wellbeing and income changed over time between 2001 and 2019. One implication of such a change is the disparity in affective wellbeing between income groups has increased. This will occur as more people fall below the change point over time and so are subject to the steep region of the function where income and affective wellbeing are strongly related. To directly examine the implications of our model for such inequities in affective wellbeing, we used the model to estimate or _predict_ the affective wellbeing of two different income levels: one income level which was above the change point in 2001 but fell below it by 2019 (\$50K/yr); and another level which always remained above the change points over the same period (\$75K/yr). The figure below presents the affective wellbeing levels for each income level as well as the difference in affective wellbeing between them (∆).

@@

##### Figure S8. Affective wellbeing at $50K/yr and $75K/yr from 2001-2019  
```{r figure_s8, fig.width=7, fig.height=5, dpi=300}
nl_fits = read_rds("../results/gh9_hhoecd_0cov_tight.RDS")

fixed_income_hash <- read_rds("../results/gh9_hhoecd_summary.RDS") %>%
  bind_rows(.id = "fit") %>%
  mutate(low = (40000 - mean)/sd,
         high = (80000 - mean)/sd) %>%
  column_to_rownames(var = "fit") %>%
  as.matrix()

@@

fixed_income_fit %>%
  mutate(year = str_remove(fit, "fit"),
         Happiness = Estimate * params_df[year, "sd_gh9"] + params_df[year, "mean_gh9"]) %>%
  select(Happiness, Dollars, year) %>%
  spread(Dollars, Happiness) %>%
  mutate(`$75K - $50K (∆)` = (`$75K` - `$50K`)) %>%
  gather(income, happiness, -year) %>%
  mutate(
    income = fct_relevel(income, "$50K", "$75K"),
    year = as.numeric(year)
    ) -> d1

d1 %>%
  filter(income %in% c("$75K", "$50K")) %>%
  ggplot(aes(x = year, y = happiness, color = income)) +
    geom_point(size = 4, alpha = 0.25) +
    geom_smooth(method = "loess", se = F, span = 3) +
    facet_wrap(~income) +
    labs(x = "", y = "") +
    ft_theme() +
    ggthemes::scale_colour_wsj() +
    theme(legend.position = "none",
          plot.title = element_text(size = 19),
          strip.text = element_text(size = 12)) -> p1

d1 %>%
  filter(!income %in% c("$75K", "$50K")) %>%
  ggplot(aes(x = year, y = happiness, color = income)) +
  geom_point(size = 4, alpha = 0.25) +
  geom_smooth(method = "loess", se = F, span = 3) +
  facet_wrap(~income) +
  labs(x = "", y = "") +
  ft_theme() +
  scale_colour_manual(values = "#be9c2e") +
  theme(legend.position = "none",
        plot.title = element_text(size = 19),
        strip.text = element_text(size = 12)) -> p2

library(patchwork)
p1 + p2 + plot_layout(widths = c(2, 1))
```

@@

The figure above shows the disparity in affective wellbeing between two fixed income levels, \$50K/year (traversing the change points) and \$75K/year (above the change points). In 2001, a household income of \$50K/year achieved an above average level of affective wellbeing relative to everyone else that year, however by 2019 affective wellbeing had declined to lower than average levels relative to the population. By comparison, a household income of \$75K/year enjoyed a higher level of affective wellbeing which decreased more slowly over the same period. The difference (∆) in affective wellbeing between these two income levels doubled over this period. The increasing disparity in affective wellbeing between the rich and the poor implies that affective wellbeing has became more inequitable over time.

@@

#### Bayesian linear, log-linear and piecewise-linear models
